---
output:
  word_document: default
  html_document: default
---

**The introduction of genomic biomarker tests for breast cancer**

Rivera Plata, Isabel;  Day, Roger

Department of Biomedical Informatics

University of Pittsburgh


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

options(width=132)
```

```{r RData}
###  CAUTION:
###   FIRST do this in the RStudio console:
###  save.image(".RData")  ---   Do NOT run this line in this document!
rm(list=ls())
attach('.RData')   ### Making sure that previously created objects are available.
```



```{r googlesheets,results='hide',message=FALSE, warnings=FALSE}
options(warn=-1)
library(googlesheets)
my_gs_meta = invisible(gs_title('Gene biomarker survey'))
my_gs = invisible(gs_read(my_gs_meta, ws = 1) [-1, ])
my_gs = invisible(data.frame(ROW=1:nrow(my_gs), my_gs))
#  Almost -- some minor but critical serror.
```


```{r excel}
library(readxl)
gbs_all = readxl::read_excel('Gene biomarker survey.xlsx')
gbs_all$ROW = 1:nrow(gbs_all)
gbs_all = gbs_all[c(ncol(gbs_all), 1:(ncol(gbs_all)-1))]
gbs = gbs_all[!is.na(gbs_all$STATUS) & (gbs_all$STATUS == 'DONE'),  ]
```


Biomarkers measure some aspects of disease to support clinical decisions. In the early 2000's, two gene expression-based biomarker tests, Oncotype DX and Mammaprint, were developed as prognostic assays for breast cancer with low risk of recurrence, to reduce unnecessary overtreatment with chemotherapy. 

We automated the extraction of `r length(titles)` relevant articles from PubMed into R via XML, did a stratified random selection of `r length(pmidSample)` articles, transferred their PubMed identifiers into a Google sheet, and developed a data dictionary. Using the sheet, we catalogued the article features of interest manually. 
Our main features of interest were the focus questions (e.g. clinical benefit vs economics vs doctor behaviors), study types (e.g. clinical, lab, or computational), and effect measures. 
We summarized the features' patterns across time and the differences between the two tests. 
We detected several examples where our personal judgments of classifications were correct in which an automated "bag-of-words" method would be incorrect.
